---
title: "Scatter Plots and Correlation: Example"
author: ""
date: ""
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
---

<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  .col3 {
    columns: 3 100px;
    -webkit-columns: 3 100px;
    -moz-columns: 3 100px;
  }
</style>

```{r setup, include=FALSE}
require(ggplot2)
require(grid)
require(dplyr)
require(tidyr)
require(readr)
```

## Today's example: Iris flower measurements

\begin{figure}[!h]
\centering
\includegraphics{Iris_germanica_wikipedia.jpg}
\caption{Iris flower (image source: Wikipedia)}
\end{figure}

#### 1. Based on the plot below, what are the observational units, variable(s), and variable type(s) in this data set?

\vspace{1cm}

#### 2. What did the code I used to make the plot look like? (note the data are in a data frame named `iris`)

\vspace{2cm}

```{r, echo=FALSE, message=FALSE, fig.height = 3.5, fig.width=6}
data(iris)
iris <- transmute(iris,
  sepal_length = Sepal.Length,
  sepal_width = Sepal.Width,
  petal_length = Petal.Length,
  petal_width = Petal.Width,
  species = Species)

ggplot() +
  geom_point(mapping = aes(x = petal_length, y = petal_width), data = iris) +
  ggtitle("Petal Length (cm) vs. Petal Width (cm) for 150 Iris Flowers")
```

\newpage

## Describe the relationship...

#### 3. Describe the shape of the relationship between these variables

\vspace{2cm}

#### 4. Are there any gaps or outliers?

\vspace{2cm}

#### 5. Would it be appropriate to use the correlation to summarize the strength and direction of the relationship between these variables?  Why or why not?

\vspace{2cm}

In case you answered "yes" to problem 5, you may use the following R output in answering the questions below:

```{r, echo = TRUE}
iris %>%
  select(petal_length, petal_width) %>%
  cor()
```

#### 6. Describe the strength of the relationship between these variables, specifically referencing the scatter plot and the correlation if appropriate.

\vspace{2cm}

#### 7. Describe the direction of the relationship between these variables, specifically referencing the scatter plot and the correlation if appropriate.

\vspace{2cm}
