---
title: ""
author: ""
date: ""
output: pdf_document
---

<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  .col3 {
    columns: 3 100px;
    -webkit-columns: 3 100px;
    -moz-columns: 3 100px;
  }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(ggplot2)
require(dplyr)
require(tidyr)
require(mvtnorm)
```

## Guess the correlation; is the correlation useful in this example?

```{r, echo=FALSE, message=FALSE}
set.seed(1)

n <- 100
mu <- c(0, 0)
sigma <- matrix(
  c(1, 0,
    0, 1),
  nrow = 2)
temp <- rmvnorm(n, mean = mu, sigma = sigma)
example <- data.frame(
  x = temp[, 1],
  y = temp[, 2]
)
```

```{r, echo = TRUE, fig.height = 6, fig.width = 6}
ggplot(data = example, mapping = aes(x = x, y = y)) +
  geom_point()
```

\newpage

```{r, echo = TRUE}
example %>%
  cor()
```

The correlation is useful in this example.

\newpage


## Guess the correlation; is the correlation useful in this example?

```{r, echo=FALSE, message=FALSE}
set.seed(1)

n <- 100
mu <- c(0, 0)
sigma <- matrix(
  c(1, 0.9,
    0.9, 1),
  nrow = 2)
temp <- rmvnorm(n, mean = mu, sigma = sigma)
example <- data.frame(
  x = temp[, 1],
  y = temp[, 2]
)
```

```{r, echo = TRUE, fig.height = 6, fig.width = 6}
ggplot(data = example, mapping = aes(x = x, y = y)) +
  geom_point()
```

\newpage

```{r, echo = TRUE}
example %>%
  cor()
```

The correlation is useful in this example.

\newpage


## Guess the correlation; is the correlation useful in this example?

```{r, echo=FALSE, message=FALSE}
set.seed(1)

n <- 100
mu <- c(0, 0)
sigma <- matrix(
  c(1, -0.6,
    -0.6, 1),
  nrow = 2)
temp <- rmvnorm(n, mean = mu, sigma = sigma)
example <- data.frame(
  x = temp[, 1],
  y = temp[, 2]
)
```

```{r, echo = TRUE, fig.height = 6, fig.width = 6}
ggplot(data = example, mapping = aes(x = x, y = y)) +
  geom_point()
```

\newpage

```{r, echo = TRUE}
example %>%
  cor()
```

The correlation is useful in this example.

\newpage


## Guess the correlation; is the correlation useful in this example?

```{r, echo=FALSE, message=FALSE}
set.seed(1)

n <- 100
mu <- c(0, 0)
sigma <- matrix(
  c(1, 0,
    0, 1),
  nrow = 2)
temp <- rmvnorm(n, mean = mu, sigma = sigma)
temp <- rbind(temp, matrix(c(10, 10), nrow = 1))

example <- data.frame(
  x = temp[, 1],
  y = temp[, 2]
)
```

```{r, echo = TRUE, fig.height = 6, fig.width = 6}
ggplot(data = example, mapping = aes(x = x, y = y)) +
  geom_point()
```

\newpage

```{r, echo = TRUE}
example %>%
  cor()
```

The correlation is not useful in this example; there is an outlier.

\newpage


## Guess the correlation

```{r, echo=FALSE, message=FALSE}
set.seed(1)

n <- 30
example <- data.frame(
  x = seq(from = -10, to = 10, length = n),
  y = seq(from = 10, to = -10, length = n)
)
```

```{r, echo = TRUE, fig.height = 6, fig.width = 6}
ggplot(data = example, mapping = aes(x = x, y = y)) +
  geom_point()
```

\newpage

```{r, echo = TRUE}
example %>%
  cor()
```

The correlation is useful in this example.

\newpage

## Guess the correlation; is the correlation useful in this example?

```{r, echo=FALSE, message=FALSE}
set.seed(1)

n <- 30
example <- data.frame(
  x = seq(from = -10, to = 10, length = n),
  y = rep(1, n)
)
```

```{r, echo = TRUE, fig.height = 6, fig.width = 6}
ggplot(data = example, mapping = aes(x = x, y = y)) +
  geom_point()
```

\newpage

```{r, echo = TRUE}
example %>%
  cor()
```

R says the correlation is `NA`, but we define the correlation to be 0.  The correlation is useful in this example.

\newpage


## Guess the correlation; is the correlation useful in this example?

```{r, echo=FALSE, message=FALSE}
set.seed(123)

n <- 100
example <- data.frame(
  x = runif(n, -10, 10)
)
example$y <- example$x^2 + rnorm(n, mean = 0, sd = 5)
```

```{r, echo = TRUE, fig.height = 6, fig.width = 6}
ggplot(data = example, mapping = aes(x = x, y = y)) +
  geom_point()
```

\newpage

```{r, echo = TRUE}
example %>%
  cor()
```

Correlation is not useful; non-linear relationship

\newpage

## Guess the correlation; is the correlation useful in this example?

```{r, echo=FALSE, message=FALSE}
set.seed(123)

n <- 300
example <- data.frame(
  x = runif(n, -20, 20)
)
example$y <- cos(example$x) + rnorm(n, 0, 0.1)
```

```{r, echo = TRUE, fig.height = 6, fig.width = 6}
ggplot(data = example, mapping = aes(x = x, y = y)) +
  geom_point()
```

\newpage

```{r, echo = TRUE}
example %>%
  cor()
```

Correlation is not useful; non-linear relationship.

\newpage

## Guess the correlation; is the correlation useful in this example?

```{r, echo=FALSE, message=FALSE}
set.seed(123)

n <- 100
example <- data.frame(
  x = runif(n, 0, 3)
)
example$y <- exp(example$x) + rnorm(n, 0, 0.1)
```

```{r, echo = TRUE, fig.height = 6, fig.width = 6}
ggplot(data = example, mapping = aes(x = x, y = y)) +
  geom_point()
```

\newpage

```{r, echo = TRUE}
example %>%
  cor()
```

Correlation is not useful; non-linear relationship
