---
title: "Scatter Plots and Correlation: Summary of Ideas"
author: ""
date: ""
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
---

<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  .col3 {
    columns: 3 100px;
    -webkit-columns: 3 100px;
    -moz-columns: 3 100px;
  }
</style>

```{r setup, include=FALSE}
require(ggplot2)
require(grid)
require(dplyr)
require(tidyr)
```


## Describing the Relationship Between Two Quantitative Variables

### A. Shape (linear or curved)

```{r shape, echo = FALSE, cache = TRUE, fig.height=1.5, fig.width = 7}
layout <- grid.layout(nrow = 1, ncol = 2, width = unit(x = rep(5, 2), units = "cm"))
grid.newpage()
pushViewport(viewport(layout = layout))

set.seed(1)

n <- 50
x1 <- runif(n, 0, 4)
df1 <- data.frame(
  x = x1,
  y = rnorm(n, mean = x1, sd = 0.1)
)

x2 <- x1
df2 <- data.frame(
  x = x2,
  y = rnorm(n, mean = 4 * exp(x2) / exp(4), sd = 0.1)
)

p1 <- ggplot(data = df1, mapping = aes(x=x, y=y)) +
  geom_point() +
  xlim(c(0, 4)) +
  ggtitle("Linear")

p2 <- ggplot(data = df2, mapping = aes(x=x, y=y)) +
  geom_point() +
  xlim(c(0, 4)) +
  ggtitle("Curved")

print(p1, vp = viewport(layout.pos.row = 1, layout.pos.col = 1))
print(p2, vp = viewport(layout.pos.row = 1, layout.pos.col = 2))
```

\vspace{-0.5cm}
### B. Are there any gaps or outliers?

```{r outliers, echo = FALSE, cache = TRUE, fig.height=1.5, fig.width = 7}
layout <- grid.layout(nrow = 1, ncol = 3, width = unit(x = rep(6, 3), units = "cm"))
grid.newpage()
pushViewport(viewport(layout = layout))

set.seed(1)

n <- 50
x1 <- runif(n, 0, 4)
df1 <- data.frame(
  x = x1,
  y = rnorm(n, mean = x1, sd = 0.1)
)

x2 <- c(runif(35, 0, 2), runif(15, 3, 4))
df2 <- data.frame(
  x = x2,
  y = rnorm(n, mean = x2, sd = 0.1)
)

x3_no <- runif(35, 0, 2)
x3 <- c(x3_no, 4, 0.5)
df3 <- data.frame(
  x = x3,
  y = c(rnorm(35, mean = x3_no, sd = 0.1), 4, 3.5),
  outlier = c(rep("Not an\noutlier", 35), "Outlier", "Outlier")
)


p1 <- ggplot(data = df1, mapping = aes(x=x, y=y)) +
  geom_point() +
  xlim(c(0, 4)) +
  ggtitle("No Gaps or Outliers")

p2 <- ggplot(data = df2, mapping = aes(x=x, y=y)) +
  geom_point() +
  xlim(c(0, 4)) +
  ggtitle("Gap")

p3 <- ggplot(data = df3, mapping = aes(x=x, y=y, color = outlier)) +
  geom_point() +
  scale_color_manual(values = c("black", "red")) +
  xlim(c(0, 4)) +
  ggtitle("Outliers") +
  theme(legend.position = "none")

print(p1, vp = viewport(layout.pos.row = 1, layout.pos.col = 1))
print(p2, vp = viewport(layout.pos.row = 1, layout.pos.col = 2))
print(p3, vp = viewport(layout.pos.row = 1, layout.pos.col = 3))
```

\vspace{-0.5cm}
### C. Direction (positive or negative association)

**Positive association**: as one variable increases, the other variable also tends to increase.

**Negative association**: as one variable increases, the other variable tends to decrease.

```{r direction, echo = FALSE, cache = TRUE, fig.height=1.5, fig.width = 7}
layout <- grid.layout(nrow = 1, ncol = 2, width = unit(x = rep(5, 2), units = "cm"))
grid.newpage()
pushViewport(viewport(layout = layout))

set.seed(1)

n <- 50
x1 <- runif(n, 0, 4)
df1 <- data.frame(
  x = x1,
  y = rnorm(n, mean = x1, sd = 0.1)
)

x2 <- x1
df2 <- data.frame(
  x = x2,
  y = rnorm(n, mean = 4 - x2, sd = 0.1)
)

p1 <- ggplot(data = df1, mapping = aes(x=x, y=y)) +
  geom_point() +
  xlim(c(0, 4)) +
  ggtitle("Positive")

p2 <- ggplot(data = df2, mapping = aes(x=x, y=y)) +
  geom_point() +
  xlim(c(0, 4)) +
  ggtitle("Negative")

print(p1, vp = viewport(layout.pos.row = 1, layout.pos.col = 1))
print(p2, vp = viewport(layout.pos.row = 1, layout.pos.col = 2))
```

\vspace{-0.5cm}
### D. Strength (weak, moderate, or strong)

The relationship is strong if the points fall close to the trend, and weak if they do not.

```{r strength, echo = FALSE, cache = TRUE, fig.height=1.5, fig.width = 7}
layout <- grid.layout(nrow = 1, ncol = 3, width = unit(x = rep(5, 3), units = "cm"))
grid.newpage()
pushViewport(viewport(layout = layout))

set.seed(1)

n <- 50
x1 <- runif(n, 0, 4)
df1 <- data.frame(
  x = x1,
  y = rnorm(n, mean = x1, sd = 0.1)
)

x2 <- x1
df2 <- data.frame(
  x = x2,
  y = rnorm(n, mean = x2, sd = 1)
)

x3 <- x1
df3 <- data.frame(
  x = x3,
  y = rnorm(n, mean = x3, sd = 2)
)

p1 <- ggplot(data = df1, mapping = aes(x=x, y=y)) +
  geom_point() +
  xlim(c(0, 4)) +
  ggtitle("Strong")

p2 <- ggplot(data = df2, mapping = aes(x=x, y=y)) +
  geom_point() +
  xlim(c(0, 4)) +
  ggtitle("Moderate")

p3 <- ggplot(data = df3, mapping = aes(x=x, y=y)) +
  geom_point() +
  xlim(c(0, 4)) +
  ggtitle("Weak")

print(p1, vp = viewport(layout.pos.row = 1, layout.pos.col = 1))
print(p2, vp = viewport(layout.pos.row = 1, layout.pos.col = 2))
print(p3, vp = viewport(layout.pos.row = 1, layout.pos.col = 3))
```


## Correlation

The correlation is **only useful if the relationship is linear and there are no outliers**.

If those conditions are met, it summarizes:

 * **Strength**:
    * Strong relationship if correlation is close to $1$ or $-1$
    * Weak relationship if correlation is close to $0$
 * **Direction**:
    * Positive association if correlation is positive
    * Negative association if correlation is negative.

## Example of Calculation in R

```{r, echo = FALSE}
data(iris)
iris <- transmute(iris,
  sepal_length = Sepal.Length,
  sepal_width = Sepal.Width,
  petal_length = Petal.Length,
  petal_width = Petal.Width,
  species = Species)
```

Suppose we want to calculate the correlation between `petal_length` and `petal_width` in the following data set with measurements on 150 iris flowers:

```{r}
head(iris)
```

We can do this with the following code:

```{r, echo = TRUE}
iris %>%
  select(petal_length, petal_width) %>%
  cor()
```

The correlation between `petal_length` and `petal_width` is about 0.963.  This indicates a strong, positive association between these variables.  We'd need to check a scatter plot to be sure it was a useful summary of the relationship though!
