---
title: "Random Variables and the Binomial Distribution"
subtitle: "(Highlights from Chapters 15 and 16)"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(ggplot2)
require(dplyr)
require(tidyr)
require(readr)
```


## Random Variables

* A **random variable** is a variable whose possible values are numerical outcomes of a random phenomenon.  Use capital letters like $X$, $Y$, $Z$ to denote random variables.  Use lower case letters $x$, $y$, $z$ to denote specific observed values.

* **Example:** $X$ = number of times Paul the Octopus correctly predicts the winner of a World Cup Soccer/Football game in 8 attempts.  He got $x = 8$ correct!

\includegraphics{paul_the_octopus.jpg}

(image credit: Wolfgang Rattay/Reuters)

* **Example:** $X$ = number of frogs in a sample of size 30 that have a certain genetic mutation.  Maybe in a particular sample I observe $x = 2$.

## Binomial Distribution

The **Binomial** distribution represents the distribution of a count of the number of "successes" in $n$ trials

 * The probability of success is the same on each trial: denote by lower case $p$
 * The trials are independent

We use a short-hand notation to describe this:

* $X \sim \text{Binomial}(n, p)$
    * "$X$ follows a Binomial distribution with $n$ trials and probability of success $p$"

\newpage

## Example in Detail: Paul the Octopus

Define $X$ = the number of successful predictions in 8 attempts.

Suppose $p = 0.8$ (Paul's predictions are pretty good!)

We could use the model $X \sim \text{Binomial}(8, 0.8)$

## Calculations with the Binomial

* Suppose $X \sim \text{Binomial}(8, 0.8)$

Define the following events (assumed to be independent):

$A_1$: Paul's 1st prediction is correct, $A_2$: Paul's 2nd prediction is correct, ..., $A_8$: Paul's 8th prediction is correct

#### a. What does the assumption that these events are independent actually mean?

\vspace{3cm}

#### b. What's the probability that Paul gets 8 out of 8 predictions correct?

```{r, echo = TRUE}
dbinom(x = 8, size = 8, prob = 0.8)
```

\vspace{3cm}

#### c. What's the probability that Paul gets the first 7 predictions correct and the last one wrong?

\vspace{3cm}

\newpage

#### d. What's the probability that Paul gets 7 out of 8 predictions correct? (we're not specifying which one he got wrong -- it could be the first, or the second, or the third, ..., or the eighth)

```{r, echo = TRUE}
dbinom(x = 7, size = 8, prob = 0.8)
```

\vspace{5cm}

#### e. What's the probabilty that Paul gets 5 or fewer predictions correct?

**Just kidding!**  Let's not calculate that by hand.

```{r, echo = TRUE}
pbinom(q = 5, size = 8, prob = 0.8)
```

\vspace{0.5cm}

## The Full Binomial Distribution

* We can use `dbinom` to calculate the probability of $x$ successes for each possible $x$ from 0 to 8:
```{r, echo = TRUE, fig.height=2}
Paul_success_probs <- data.frame(
  num_successes = seq(from = 0, to = 8),
  probability = dbinom(x = seq(from = 0, to = 8), size = 8, prob = 0.9))
Paul_success_probs
ggplot() +
  geom_col(mapping = aes(x = num_successes, y = probability),
    data = Paul_success_probs)
```

## Expected Value

* How many do we **expect** Paul to get right?  (same as **average**, or **mean**)
* Weighted average of the possible values of $X$, weighted according to how likely they are

\begin{align*}
&\mu = E(X) = \sum_x x P(X = x)
\end{align*}

\vspace{4cm}


## Variance and Standard Deviation

* Variance: $\sigma^2 = Var(X) = \sum_x (x - \mu)^2 P(X = x)$

\vspace{5cm}

* Standard Deviation: $\sigma = SD(X) = \sqrt{Var(X)}$
