---
title: "Stat 140: More Probability Examples"
output:
  pdf_document:
    fig_height: 2.8
    fig_width: 6
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


```{r include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
require(readr)
require(ggplot2)
require(dplyr)
```

```{r include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

# Diabetes and hypertension

Diabetes and hypertension are two of the most common diseases in Western, industrialized nations. In the United States, approximately 9\% of the population have diabetes, while about 30\% of adults have high blood pressure. The two diseases frequently occur together: an estimated 6\% of the population have both diabetes and hypertension.


### 1. Are having diabetes and having hypertension disjoint?

\vspace{1.5cm}

### 2. Draw a Venn diagram summarizing the variables and their associated probabilities.

\vspace{4cm}

### 3. Let $A$ represent the event of having diabetes, and $B$ the event of having hypertension. Calculate $P(A \text{ or } B)$.

\vspace{4cm}

### 4. What percent of Americans have neither hypertension nor diabetes?

\newpage

# Drug Testing

Mandatory drug testing in the workplace is common practice for certain professions, such as air traffic controllers and transportation workers.  A false positive in a drug screening test occurs when the test incorrectly indicates that a screened person is an illegal drug user.  Suppose a mandatory drug test has a false positive rate of 1.2\% (i.e., has probability  0.012 of indicating that an employee is using illegal drugs when that is not the case).

### 1. A company gives all 150 of its employees drug tests, and in reality all 150 employees are not drug users.  Assume that the drug tests are \textbf{independent}: the outcome of one drug test has no effect on the others.  Define the events:

\begin{align*}
A_1 &= \text{ the 1st employee tests positive} \\
A_2 &= \text{ the 2nd employee tests positive} \\
 &\vdots \\
A_{150} &= \text{ the 150th employee tests positive} \\
\end{align*}


#### a. Find the probability that a particular employee tests negative

\vspace{1.5cm}

#### b. Find the probability that all 150 employees test negative

\vspace{3cm}

#### c. Find the probability that at least one employee (falsely) tests positive

\vspace{3cm}

\newpage

### 2.

Because of the high likelihood of at least one false positive in company wide drug screening programs, an individual with a positive test is almost always re-tested with a different screening test: one that is more expensive than the first, but has a lower false positive probability. Suppose the second test has a false positive rate of 0.8\%.

What is the probability that an employee who is not using illegal drugs will test positive on both tests?  Assume the two tests are independent of each other.


