---
title: "Stat 140: Probability Examples"
output:
  pdf_document:
    fig_height: 2.8
    fig_width: 6
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


```{r include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
require(readr)
require(ggplot2)
require(dplyr)
```

```{r include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

# Blood Types

The American Red Cross says that about 45% of the US Population has Type O blood, 40% has Type A blood, 11% has Type B blood, and the rest have Type AB blood.  (A single person will have one blood type -- I can't have both Type A and Type B blood.)

### 1. Someone volunteers to give blood.

#### (a) Are the events that this donor has type A blood and that they have type B blood disjoint?

\vspace{3cm}

#### (b) What is the probability that this donor has Type A or Type B blood?

\vspace{3cm}

#### (c) What is the probability that this donor has Type A blood, Type B blood, or Type O blood?

\vspace{3cm}

#### (d) What is the probability that this donor has Type AB blood? (Use the complement rule).

\vspace{3cm}

### 2. Four potential blood donors come into to the clinic.  They are not related, so we can treat their blood types as being independent.

#### (a) What is the probability that all four are Type O?

\vspace{3cm}

#### (b) What is the probability that they are not all Type O?  (That is, what is the probability that at least one of them has a blood type other than O).  (Use the complement rule).

\vspace{3cm}

#### (c) What is the probability that no one is Type AB?

\vspace{3cm}

\newpage

# Holiday Meals

A survey of American families revealed that 58\% of all families eat turkey at holiday meals, 44\% eat ham, and 16\% have both turkey and ham to eat at holiday meals.

Suppose we pick a family at random.  Define the events

A = the family eats turkey at holiday meals

B = the family eats ham at holiday meals

#### (a) What is P(A)?  What is P(B)?  What is P(A and B)?  

\vspace{3cm}

#### (b) What is the probability that a family selected at random had turkey or ham (or both) at their holiday meal?

\vspace{3cm}

#### (c) What is the probability that a family selected at random had neither turkey nor ham at their holiday meal?

\vspace{3cm}

#### (d) What is the probability that a family selected at random had only ham without having turkey at their holiday meal?  (Draw a diagram)

\vspace{3cm}

#### (e) Are having turkey and having ham disjoint events? Explain.\\

\vspace{3cm}
