---
title: 'Stat 140: Rules for Probabilities'
output:
  pdf_document:
    fig_height: 2.9
    fig_width: 6.5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
require(tidyr)
options(digits=3)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
# This loads the mosaic data sets.  (Could be deleted if you are not using them.)
require(mosaicData)                
```

# Probability (Chapters 13 and 14)

## General Terminology

1. Trial: a single attempt or realization of a random phenomenon

2. Outcome: an observed value in a trial

3. Event: A set of possible outcomes.  Use letters like $A$, $B$, $C$

4. Sample Space: The set of all possible outcomes.  Use $S$.

5. Complement: The complement of the event $A$ is the set of all possible outcomes *not* in $A$.  Use $A^C$.

6. Disjoint Events: have no outcomes in common

7. Independent Events: knowing one occurred doesn't change what you know about the chances of the other occurring.

## Probability Definitions and Rules

Let $A$ and $B$ be events, and let $S$ be the sample space.

#### Foundations

* $0 \leq P(A) \leq 1$

* $P(S) = 1$

* $P(A^c) = 1 - P(A)$

#### Probability of A *or* B (or both) occurring

* $P(A\text{ or }B) = P(A) + P(B) - P(A\text{ and }B)$

* If $A$ and $B$ are disjoint events, then $P(A\text{ or }B) = P(A) + P(B)$.

#### Conditional Probability; Probability of A *and* B both occurring

* Conditional Probability of $B$ given $A$: $P(B | A) = \frac{P(\text{B and } A)}{P(A)}$

* $P(A \text{ and } B) = P(A) \times P(B | A) = P(B) \times P(A | B)$

* If $A$ and $B$ are **independent** events, then $P(A \text{ and } B) = P(A) \times P(B)$.

* $A$ and $B$ are **independent** if (and only if) $P(B | A) = P(B)$, or turning that around, $P(A | B) = P(A)$.

#### Bayes' Rule

$P(A | B) = \frac{P(B | A)P(A)}{P(B | A)P(A) + P(B | A^c)P(A^c)}$
