---
title: "Boxplots, histograms, and density plots, oh my!"
author: ''
date: "February 6, 2018"
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
library(ggplot2)
require(mosaic)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
# This loads the mosaic data sets.  (Could be deleted if you are not using them.)
require(mosaicData)                
```

```{r, echo=FALSE}
## Create a fake data set.  It's a secret at first, but the data have the same 5 number summary!
## I got this code from Stack Overflow: see the comment by Glen_b at
## http://stats.stackexchange.com/questions/137965/box-and-whisker-plot-for-multimodal-distribution
set.seed(1)

x1 = qnorm(ppoints(1:100,a=-.072377))
x1 = x1/diff(range(x1))*18+10
b = fivenum(x1)
x2 = qnorm(ppoints(1:48));x2=x2/diff(range(x2))*.6
x2 = c(b[1],x2+b[2],.31+b[2],b[4]-.31,x2+b[4],b[5])
d = .1183675; x3 = ((0:34)-34/2)/34*(9-d)+(5.5-d/2)
x3 = c(x3,rep(9.5,15),rep(10.5,15),20-x3)
x4 = c(1,rep(b[2],24),(0:49)/49*(b[4]-b[2])+b[2],(0:24)/24*(b[5]-b[4])+b[4])

example_data <- data.frame(
  x1 = x1,
  x2 = x2,
  x3 = x3,
  x4 = x4
)
```

Here are some box (and whiskers) plots of four different data sets:

```{r, fig.height = 1.5}
ggplot(data = example_data, mapping = aes(x = "Data Set 1", y = x1)) + geom_boxplot()
```

```{r, fig.height = 1.5}
ggplot(data = example_data, mapping = aes(x = "Data Set 2", y = x2)) + geom_boxplot()
```

```{r, fig.height = 1.5}
ggplot(data = example_data, mapping = aes(x = "Data Set 3", y = x3)) + geom_boxplot()
```

```{r, fig.height = 1.5}
ggplot(data = example_data, mapping = aes(x = "Data Set 4", y = x4)) + geom_boxplot()
```

What can you tell me about the distributions based on these box plots?  **Shape**?  **Outliers**?  Anything else?

\newpage

Here are some histograms of the same data:

```{r, fig.height = 1.5}
ggplot(data = example_data, mapping = aes(x = x1)) + geom_histogram(binwidth = 1)
```

```{r, fig.height = 1.5}
ggplot(data = example_data, mapping = aes(x = x2)) + geom_histogram(binwidth = 1)
```

```{r, fig.height = 1.5}
ggplot(data = example_data, mapping = aes(x = x3)) + geom_histogram(binwidth = 1)
```

```{r, fig.height = 1.5}
ggplot(data = example_data, mapping = aes(x = x4)) + geom_histogram(binwidth = 1)
```

What can you tell me about the distributions based on these histograms?  **shape**?  **outliers**?  Anything else?


\newpage

Here are some density plots of the same data:

```{r, fig.height = 1.5}
ggplot(data = example_data, mapping = aes(x = x1)) + geom_density()
```

```{r, fig.height = 1.5}
ggplot(data = example_data, mapping = aes(x = x2)) + geom_density()
```

```{r, fig.height = 1.5}
ggplot(data = example_data, mapping = aes(x = x3)) + geom_density()
```

```{r, fig.height = 1.5}
ggplot(data = example_data, mapping = aes(x = x4)) + geom_density()
```

What can you tell me about the distributions based on these density plots?  **shape**?  **outliers**?  Anything else?
