---
title: 'Stat 140: Inference for Simple Linear Regression Example - Wild Horses'
author: "Evan Ray"
output:
  pdf_document:
    fig_height: 2.7
    fig_width: 6.5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
header-includes:
   - \usepackage{multicol}
---


```{r, include=FALSE}
require(readr)
require(tidyr)
require(dplyr)
require(ggplot2)
require(knitr)
require(mvtnorm)

options(digits=4)

opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

# Wild Horses

What is the relationship between the size of a herd of horses and the number of foals (baby horses!!) that are born to that herd in a year?

```{r}
horses <- read_csv("https://mhc-stat140-2017.github.io/data/sdm4/Wild_Horses.csv")
head(horses)
nrow(horses)
```

## Questions to Start With:

 * What is the observational unit?

\vspace{0.5cm}

 * What are the variable data types (**categorical** or **quantitative**)?
     * **Foals**: 
     * **Adults**: 

\vspace{0.5cm}

 * Which of these variables is the **explanatory** variable and which is the **response**?
     * **Explanatory**: 
     * **Response**: 

\newpage

## Previously: Fit linear regression to describe the relationship between number of adults and number of foals in the *sample*.

 * Estimated Line: $\hat{y}_i = b_0 + b_1 x_i$
 * $b_0$ and $b_1$ are **sample statistics**: they describe the data in our sample

```{r, echo = FALSE}
ggplot() +
  geom_point(mapping = aes(x = Adults, y = Foals), color = "red", data = horses) +
  geom_smooth(mapping = aes(x = Adults, y = Foals),
    method = "lm",
    color = "red",
    se = FALSE,
    data = horses) +
  xlim(c(20, 280)) +
  ylim(c(-5, 50))
```

## Today: Use data from this sample to learn about the relationship between number of adults and number of foals in the *population*

 * Population Model:
    * $y_i = \beta_0 + \beta_1 x_i + \varepsilon_i$
    * $\varepsilon_i \sim \text{Normal}(0, \sigma)$
 * $\beta_0$ and $\beta_1$ are **population parameters**: they describe the population

```{r, echo = FALSE}
lm_fit <- lm(Foals ~ Adults, data = horses)

N <- 3 * nrow(horses)

set.seed(1)
population_coeffs <- c(-1, 0.138)

horses_population <- data.frame(
  Adults = runif(n = N, 25, 270)
)
horses_population$Foals <-
  population_coeffs[1] + population_coeffs[2] * horses_population$Adults + rnorm(N, mean = 0, sd = summary(lm_fit)$sigma)
horses_population$In_Sample = FALSE
horses_population <- bind_rows(horses_population,
  transmute(horses,
    Adults = Adults,
    Foals = Foals,
    In_Sample = TRUE)
)

ggplot() +
  geom_point(mapping = aes(x = Adults, y = Foals), data = horses_population) +
  geom_smooth(mapping = aes(x = Adults, y = Foals),
    method = "lm",
    color = "black",
    se = FALSE,
    data = horses_population) +
  geom_smooth(mapping = aes(x = Adults, y = Foals),
    method = "lm",
    color = "red",
    se = FALSE,
    data = horses) +
  xlim(c(20, 280)) +
  ylim(c(-5, 50))
```

\newpage

\includegraphics{pop-sample.png}

\newpage

## (a) Are the assumptions for the linear regression model met?

\includegraphics[width=0.5\textwidth]{leprechaun.png}


\begin{itemize}
\item \textbf{Representative} Sample
\vspace{2cm}

\item (No) \textbf{Outliers}
\vspace{2cm}

\item \textbf{Linear} Relationahip (Straight Enough)
\vspace{2cm}

\item \textbf{Independent} Observations (Randomization)
\vspace{2cm}

\item \textbf{Normal} Distribution of Residuals (Can't check this yet -- need to look at a histogram or density plot of the residuals after fitting the model)
\vspace{2cm}

\item \textbf{Equal} Variance of Residuals (Does the Plot Thicken?)
\end{itemize}



\vspace{4cm}

\newpage

## (b) Fit the linear model

```{r}
# format is: lm(response_variable ~ explanatory_variable, data = data_frame)
lm_fit <- lm(Foals ~ Adults, data = horses)
summary(lm_fit)
```

## (c) Check that the residuals follow a nearly normal distribution

```{r}
horses <- mutate(horses,
  residual = residuals(lm_fit),
  predicted = predict(lm_fit))
ggplot() +
  geom_density(mapping = aes(x = residual), data = horses)
```

## (d) Explain in context what the regression says about the relationship between the number of adult horses in a herd and the number of foals born to that herd.  Interpret both the intercept and the slope in context.

\vspace{7cm}

## (e) Conduct a hypothesis test of the claim that when there are 0 adults in a herd, there will be 0 foals born to that herd.

\vspace{4cm}

## (f) Draw a picture of a relevant t distribution for the hypothesis test in part (e) and shade in the region corresponding to the p-value.  How would you calculate the p-value for part (e) using the pt function in R and the given estimate and standard error?

\vspace{4cm}

\newpage

## (g) Conduct a hypothesis test of the claim that there is no relationship between the number of adults in a herd and the number of foals who are born to that herd.

\vspace{4cm}

## (h) Obtain a 99% confidence interval for the population intercept, $\beta_0$, and for the population slope, $\beta_1$.  Interpret the confidence interval for $\beta_1$ in context.

```{r}
## Note that unlike every other confidence interval function we've looked at,
## we set the confidence level with an argument called level, not conf.level
confint(lm_fit, level = 0.99)
```

\vspace{4cm}

## (i) How would you calculate the confidence interval for part (f) using the `qt` function in R and the given estimate and standard error?

```{r}
qt(0.995, df = 38 - 2)
```

\vspace{3cm}

## (j) Interpret the standard error for the slope using the "95" part of the 68-95-99.7 rule.

