---
title: "Linear Regression - First Example"
output: pdf_document
header-includes:
   - \usepackage{soul}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
library(readr)
library(dplyr)
library(ggplot2)
library(mosaic)
```

# Mortality and Hard Water

* Scientists believe that hard water (water with high concentrations of calcium and magnesium) is beneficial for health (Sengupta, P. (2013). IJPM, 4(8), 866-875.)

* We have recordings of the mortality rate (deaths per 100,000 population) and concentration of calcium in drinking water (parts per million) in 61 large towns in England and Wales

```{r, message=FALSE, fig.height = 5, fig.width=6}
mortality_water <- read_csv("https://mhc-stat140-2017.github.io/data/sdm4/Hard_water_Derby.csv")

ggplot(data = mortality_water, mapping = aes(x = Calcium, y = Mortality)) +
  geom_point()
```

* **Explanatory** (~~independent~~) variable goes on the x axis, **Response** (~~dependent~~) variable on the y axis
* **Notation**:
    * $x_i$: value of explanatory variable (Calcium) for observational unit number $i$
    * $y_i$: value of response variable (Mortality) for observational unit number $i$
* **Big idea**: Summarize the relationship between these variables with a line.

\newpage

# R Commands (first look - more to come)

## Add a line to the plot

```{r, echo=TRUE, message=FALSE, fig.height = 2, fig.width=6}
ggplot(data = mortality_water, mapping = aes(x = Calcium, y = Mortality)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

## Estimate intercept and slope of line

```{r}
linear_fit <- lm(Mortality ~ Calcium, data = mortality_water)
```

General form of `lm` command:

```{r, echo = TRUE, eval = FALSE}
lm(response_variable ~ explanatory_variable, data = data_frame)
```

## View summary of linear model fit

```{r}
summary(linear_fit)
```

\newpage

## Foundations

### 1. What is the estimated intercept and its interpretation?

\vspace{3cm}

### 2. What is the estimated slope and its interpretation?

\vspace{3cm}

### 3. What is the estimated equation for the regression line?

\vspace{3cm}

### 4. One of the towns in our sample had a measured Calcium concentration of 71.  What is the predicted value for the mortality rate in that town?
```{r}
predict_data <- data.frame(
  Calcium = 71
)
predict(linear_fit, newdata = predict_data)
```

\vspace{3cm}

### 5. Based on this analysis, does increasing the concentration of Calcium in drinking water cause a reduction in the mortality rate?

\newpage

## Residuals

* $\definecolor{residual}{RGB}{230,159,0}\color{residual}\text{Residual}$ = $\definecolor{observed}{RGB}{0,158,115}\color{observed}Observed$ - $\definecolor{predicted}{RGB}{86,180,233}\color{predicted}Predicted$

* $\definecolor{residual}{RGB}{230,159,0}\color{residual}e_i$ = $\definecolor{observed}{RGB}{0,158,115}\color{observed}y_i$ - $\definecolor{predicted}{RGB}{86,180,233}\color{predicted}\widehat{y}_i$ ($e$ stands for error)

```{r, echo=FALSE, message=FALSE, fig.height = 4, fig.width=6}
x <- 71
y_obs <- 1713
y_hat <- predict(linear_fit, newdata = data.frame(Calcium = 71))

ex_df <- data.frame(
  x = x,
  y_obs = y_obs,
  y_hat = y_hat,
  resid = y_obs - y_hat
)

offset <- 8
offset2 <- 5
y_mid <- mean(c(y_obs, y_hat))
resid_df <- data.frame(
  x = c(x, x + offset, x + offset, x + offset + offset2, x + offset, x + offset, x),
  y = c(y_obs, y_obs, y_mid, y_mid, y_mid, y_hat, y_hat)
)

ggplot() +
  geom_point(mapping = aes(x = Calcium, y = Mortality), data = mortality_water) +
  geom_smooth(mapping = aes(x = Calcium, y = Mortality), data = mortality_water, method = "lm", se = FALSE) +
  geom_path(mapping = aes(x = x, y = y), color = "#e69f00", size = 2, data = resid_df) +
  geom_point(mapping = aes(x = x, y = y_obs), color = "#009E73", size = 4, data = ex_df) +
  geom_point(mapping = aes(x = x, y = y_hat), color = "#56B4E9", size = 4, data = ex_df) +
  geom_label(mapping = aes(label = "y[i]", x = x, y = y_obs), size = 8, nudge_x = -6.5, nudge_y = 40, color = "#009E73", parse = TRUE, data = ex_df) +
  geom_label(mapping = aes(label = "hat(y)[i]", x = x, y = y_hat), size = 8, nudge_x = -6.5, nudge_y = 55, color = "#56B4E9", parse = TRUE, data = ex_df) +
  geom_label(mapping = aes(label = "e[i]", x = x + offset + offset2, y = y_mid), size = 8, nudge_x = 5, color = "#e69f00", parse = TRUE, data = ex_df)
```

### 6. For the town we considered in part 4 above, with an observed Calcium measurement of 71 ppm, the observed Mortality rate was 1713 deaths per 100,000 population.  What was the residual for that town?

\vspace{3cm}

### 7. Another town had an observed Calcium measurement of 50 ppm, and an observed Mortality rate of 1257 deaths per 100,000 population.  What was the residual for that town?
```{r}
predict(linear_fit, newdata = data.frame(Calcium = 50))
```

\newpage

## Finding the Line of Best Fit

```{r, echo=FALSE, message=FALSE, fig.height = 3, fig.width=6}
mortality_water <- mutate(mortality_water,
  Mortality = as.numeric(Mortality),
  predicted = predict(linear_fit),
  residual = residuals(linear_fit))
mortality_water_augmented <- mortality_water
mortality_water_augmented$min_val <- as.numeric(apply(mortality_water_augmented, 1, function(rowvals) {min(rowvals[c("predicted", "Mortality")])}))
mortality_water_augmented$max_val <- as.numeric(apply(mortality_water_augmented, 1, function(rowvals) {max(rowvals[c("predicted", "Mortality")])}))

ggplot() +
  geom_linerange(mapping = aes(x = Calcium, ymin = min_val, ymax = max_val), color = "#e69f00", size = 2, data = mortality_water_augmented) +
  geom_point(mapping = aes(x = Calcium, y = Mortality), color = "#009E73", size = 3, data = mortality_water_augmented) +
  geom_smooth(mapping = aes(x = Calcium, y = Mortality), data = mortality_water_augmented, method = "lm", se = FALSE) +
  geom_point(mapping = aes(x = Calcium, y = predicted), color = "#56B4E9", size = 3, data = mortality_water_augmented)
```

* The "best" line has the smallest residuals
* Pick $b_0, b_1$ to minimize sum of squared errors/residuals: $SSE = \sum_{i = 1}^n e_i^2$

$$b_1 = r \frac{s_y}{s_x} \qquad \qquad b_0 = \bar{y} - b_1 \bar{x}$$

## How useful is the line?

Rephrased: How close to the line do the points fall?

### Answer 1: $R^2$

Recall: Correlation ($r$) is a measure of the strength of a linear relationship.

$R^2$ (read as "R squared") is the square of the correlation.

"Proportion of variability in the response variable accounted for by the linear model."  (See office hours or intermediate statistics for more.)

### 8. What is the $R^2$ value for this model fit?

```{r}
mortality_water %>%
  select(Calcium, Mortality) %>%
  cor()
(-0.6548486)^2
```


\newpage

### Answer 2: How big do the residuals tend to be?

```{r, fig.height = 2}
mortality_water <- mortality_water %>% mutate(
  predicted = predict(linear_fit),
  residual = residuals(linear_fit)
)
head(mortality_water)

ggplot(data = mortality_water, mapping = aes(x = residual)) + geom_histogram(bins = 12)
```

If the residuals are all close to 0, the observed values are all close to the predicted values (i.e., the points are close to the line).

### 9. What is the residual standard deviation?

\vspace{2cm}

### 10. Assuming that the residuals follow an approximately normal distribution, what is the interpretation of the residual standard deviation based on the "95" part of the 68-95-99.7 rule?
