---
title: "Linear Regression Lab Wrap-Up"
author: "Evan L. Ray"
date: "April 13, 2018"
output: ioslides_presentation
---

<script type="text/x-mathjax-config"> 
    MathJax.Hub.Config({ 
        "HTML-CSS": { scale: 90, linebreaks: { automatic: true } }, 
        SVG: { linebreaks: { automatic:true } }, 
        displayAlign: "left" });
</script>

<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  .col3 {
    columns: 3 100px;
    -webkit-columns: 3 100px;
    -moz-columns: 3 100px;
  }
  .MathJax {
    font-size: 0.9em;
  }
</style>

<script type="text/x-mathjax-config">
  MathJax.Hub.Config({ TeX: { extensions: ["color.js"] }});
</script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(ggplot2)
require(dplyr)
require(tidyr)
require(readr)
```

## Wrap Up for Lab

```{r, message=FALSE, echo = FALSE}
wildfires <- read_csv("https://mhc-stat140-2017.github.io/data/sdm4/Wildfires_2012.csv")
names(wildfires) <- c("num_fires", "years_since_1985", "ave_acres_burned")
wildfires <- mutate(wildfires, year = years_since_1985 + 1985)
wildfires <- arrange(wildfires, year)
```

```{r, echo = TRUE}
linear_fit <- lm(ave_acres_burned ~ years_since_1985,
  data = wildfires)
coef(linear_fit)
```
* Predicted average number of acres burned in 1985 is 19.6

```{r, echo = TRUE}
linear_fit_year <- lm(ave_acres_burned ~ year,
  data = wildfires)
coef(linear_fit_year)
```
* Predicted average number of acres burned in year 0 is -4390

## What's going on?

<div class="col2">

```{r, echo = FALSE, fig.width=3.5, fig.height=4}
ggplot() +
  geom_point(aes(x = years_since_1985, y = ave_acres_burned), data = wildfires) +
  geom_abline(intercept = coef(linear_fit)[1], slope = coef(linear_fit)[2]) +
  coord_cartesian(xlim = range(wildfires$years_since_1985), ylim = c(coef(linear_fit)[1], 110))
```


```{r, echo = FALSE, fig.width=3.5, fig.height=4}
ggplot() +
  geom_point(aes(x = year, y = ave_acres_burned), data = wildfires) +
  geom_abline(intercept = coef(linear_fit_year)[1], slope = coef(linear_fit_year)[2]) +
  coord_cartesian(xlim = c(0, 2017), ylim = c(coef(linear_fit_year)[1], 110))
```
</div>

## Never Extrapolate Beyond the Data

[An Important Message](https://www.causeweb.org/cause/sites/default/files/resources/fun/videos/How%20Far%20He%E2%80%99ll%20Go%20-%20Final.mp4)
