---
title: "Linear Model Theory (ish)"
author: "Evan L. Ray"
output:
  ioslides_presentation:
    smaller: true
---

<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  .col3 {
    columns: 3 100px;
    -webkit-columns: 3 100px;
    -moz-columns: 3 100px;
  }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
require(ggplot2)
require(scales)
require(dplyr)
require(tidyr)
require(readr)
require(mosaic)
require(nycflights13)
```

## Population: Every Flight that Departed from NYC in 2013

```{r, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE, fig.height=3}
ggplot(data = flights, mapping = aes(x = distance,  y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Population Intercept: ", beta[0], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[1], digits = 4)), ", Population Slope: ", beta[1], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[2], digits = 4)))))

set.seed(2)
```

Population Model:

$Y_i = \beta_0 + \beta_1 X_i + \varepsilon_i$

$\varepsilon_i \sim \text{Normal}(0, \sigma)$

## Residuals Distribution in Population:

Not exactly normal, but close enough.

Black: Actual distribution of residuals; Red: The closest normal approximation

```{r, cache = TRUE, fig.height=3}
pop_lm <- lm(air_time ~ distance, data = flights)

pop_resids <- residuals(pop_lm)
ggplot(mapping = aes(x = pop_resids)) +
  geom_density() +
  stat_function(fun = dnorm, args = list(mean = 0, sd = sd(pop_resids)), color = "red")
```

## Sampling Distribution of $b_1$

The distribution of slope estimates $b_1$, across all different samples

```{r, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE, fig.height=1.5}
ggplot(data = flights, mapping = aes(x = distance,  y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Population Intercept: ", beta[0], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[1], digits = 4)), ", Population Slope: ", beta[1], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[2], digits = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height=1.5, warning = FALSE, eval = TRUE}
n <- 20
flights_sample <- flights %>% sample_n(size = n)
sample_lm <- lm(air_time ~ distance, data = flights_sample)
sample_slopes <- data.frame(
  sample_slope = coef(sample_lm)[2]
)
ggplot(data = flights_sample, mapping = aes(x = distance, y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Sample Intercept: ", b[0], " = ", .(round(coef(sample_lm)[1], digits = 4)), ", Sample Slope: ", b[1], " = ", .(round(coef(sample_lm)[2], digits = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height = 1.25, eval = TRUE}
ggplot() +
  geom_histogram(mapping = aes(x = sample_slope, y = ..density..), binwidth = 0.001, data = sample_slopes) +
  xlab("Sample Slopes, n = 20") +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm")) +
  xlim(c(0.1, 0.16))
```

## Sampling Distribution of $b_1$

The distribution of slope estimates $b_1$, across all different samples

```{r, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE, fig.height=1.5}
ggplot(data = flights, mapping = aes(x = distance,  y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Population Intercept: ", beta[0], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[1], digits = 4)), ", Population Slope: ", beta[1], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[2], digits = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height=1.5, warning = FALSE, eval = TRUE}
n <- 20
flights_sample <- flights %>% sample_n(size = n)
sample_lm <- lm(air_time ~ distance, data = flights_sample)
sample_slopes <- bind_rows(
  sample_slopes,
  data.frame(
    sample_slope = coef(sample_lm)[2]
  )
)

ggplot(data = flights_sample, mapping = aes(x = distance, y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Sample Intercept: ", b[0], " = ", .(format(round(coef(sample_lm)[1], digits = 4), nsmall = 4)), ", Sample Slope: ", b[1], " = ", .(format(round(coef(sample_lm)[2], digits = 4), nsmall = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height = 1.25, eval = TRUE}
ggplot() +
  geom_histogram(mapping = aes(x = sample_slope, y = ..density..), binwidth = 0.001, data = sample_slopes) +
  xlab("Sample Slopes, n = 20") +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm")) +
  xlim(c(0.1, 0.16))
```

## Sampling Distribution of $b_1$

The distribution of slope estimates $b_1$, across all different samples

```{r, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE, fig.height=1.5}
ggplot(data = flights, mapping = aes(x = distance,  y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Population Intercept: ", beta[0], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[1], digits = 4)), ", Population Slope: ", beta[1], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[2], digits = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height=1.5, warning = FALSE, eval = TRUE}
n <- 20
flights_sample <- flights %>% sample_n(size = n)
sample_lm <- lm(air_time ~ distance, data = flights_sample)
sample_slopes <- bind_rows(
  sample_slopes,
  data.frame(
    sample_slope = coef(sample_lm)[2]
  )
)

ggplot(data = flights_sample, mapping = aes(x = distance, y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Sample Intercept: ", b[0], " = ", .(format(round(coef(sample_lm)[1], digits = 4), nsmall = 4)), ", Sample Slope: ", b[1], " = ", .(format(round(coef(sample_lm)[2], digits = 4), nsmall = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height = 1.25, eval = TRUE}
ggplot() +
  geom_histogram(mapping = aes(x = sample_slope, y = ..density..), binwidth = 0.001, data = sample_slopes) +
  xlab("Sample Slopes, n = 20") +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm")) +
  xlim(c(0.1, 0.16))
```

## Sampling Distribution of $b_1$

The distribution of slope estimates $b_1$, across all different samples

```{r, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE, fig.height=1.5}
ggplot(data = flights, mapping = aes(x = distance,  y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Population Intercept: ", beta[0], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[1], digits = 4)), ", Population Slope: ", beta[1], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[2], digits = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height=1.5, warning = FALSE, eval = TRUE}
n <- 20
flights_sample <- flights %>% sample_n(size = n)
sample_lm <- lm(air_time ~ distance, data = flights_sample)
sample_slopes <- bind_rows(
  sample_slopes,
  data.frame(
    sample_slope = coef(sample_lm)[2]
  )
)

ggplot(data = flights_sample, mapping = aes(x = distance, y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Sample Intercept: ", b[0], " = ", .(format(round(coef(sample_lm)[1], digits = 4), nsmall = 4)), ", Sample Slope: ", b[1], " = ", .(format(round(coef(sample_lm)[2], digits = 4), nsmall = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height = 1.25, eval = TRUE}
ggplot() +
  geom_histogram(mapping = aes(x = sample_slope, y = ..density..), binwidth = 0.001, data = sample_slopes) +
  xlab("Sample Slopes, n = 20") +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm")) +
  xlim(c(0.1, 0.16))
```

## Sampling Distribution of $b_1$

The distribution of slope estimates $b_1$, across all different samples

```{r, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE, fig.height=1.5}
ggplot(data = flights, mapping = aes(x = distance,  y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Population Intercept: ", beta[0], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[1], digits = 4)), ", Population Slope: ", beta[1], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[2], digits = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height=1.5, warning = FALSE, eval = TRUE}
n <- 20
flights_sample <- flights %>% sample_n(size = n)
sample_lm <- lm(air_time ~ distance, data = flights_sample)
sample_slopes <- bind_rows(
  sample_slopes,
  data.frame(
    sample_slope = coef(sample_lm)[2]
  )
)

ggplot(data = flights_sample, mapping = aes(x = distance, y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Sample Intercept: ", b[0], " = ", .(format(round(coef(sample_lm)[1], digits = 4), nsmall = 4)), ", Sample Slope: ", b[1], " = ", .(format(round(coef(sample_lm)[2], digits = 4), nsmall = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height = 1.25, eval = TRUE}
ggplot() +
  geom_histogram(mapping = aes(x = sample_slope, y = ..density..), binwidth = 0.001, data = sample_slopes) +
  xlab("Sample Slopes, n = 20") +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm")) +
  xlim(c(0.1, 0.16))
```

## Sampling Distribution of $b_1$

The distribution of slope estimates $b_1$, across all different samples

```{r, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE, fig.height=1.5}
ggplot(data = flights, mapping = aes(x = distance,  y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Population Intercept: ", beta[0], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[1], digits = 4)), ", Population Slope: ", beta[1], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[2], digits = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height=1.5, warning = FALSE, eval = TRUE}
n <- 20
flights_sample <- flights %>% sample_n(size = n)
sample_lm <- lm(air_time ~ distance, data = flights_sample)
sample_slopes <- bind_rows(
  sample_slopes,
  data.frame(
    sample_slope = coef(sample_lm)[2]
  )
)

ggplot(data = flights_sample, mapping = aes(x = distance, y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Sample Intercept: ", b[0], " = ", .(format(round(coef(sample_lm)[1], digits = 4), nsmall = 4)), ", Sample Slope: ", b[1], " = ", .(format(round(coef(sample_lm)[2], digits = 4), nsmall = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height = 1.25, eval = TRUE}
ggplot() +
  geom_histogram(mapping = aes(x = sample_slope, y = ..density..), binwidth = 0.001, data = sample_slopes) +
  xlab("Sample Slopes, n = 20") +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm")) +
  xlim(c(0.1, 0.16))
```

## Sampling Distribution of $b_1$

The distribution of slope estimates $b_1$, across all different samples

```{r, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE, fig.height=1.5}
ggplot(data = flights, mapping = aes(x = distance,  y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Population Intercept: ", beta[0], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[1], digits = 4)), ", Population Slope: ", beta[1], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[2], digits = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height=1.5, warning = FALSE, eval = TRUE}
n <- 20
flights_sample <- flights %>% sample_n(size = n)
sample_lm <- lm(air_time ~ distance, data = flights_sample)
sample_slopes <- bind_rows(
  sample_slopes,
  data.frame(
    sample_slope = coef(sample_lm)[2]
  )
)

ggplot(data = flights_sample, mapping = aes(x = distance, y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Sample Intercept: ", b[0], " = ", .(format(round(coef(sample_lm)[1], digits = 4), nsmall = 4)), ", Sample Slope: ", b[1], " = ", .(format(round(coef(sample_lm)[2], digits = 4), nsmall = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height = 1.25, eval = TRUE}
ggplot() +
  geom_histogram(mapping = aes(x = sample_slope, y = ..density..), binwidth = 0.001, data = sample_slopes) +
  xlab("Sample Slopes, n = 20") +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm")) +
  xlim(c(0.1, 0.16))
```

## Sampling Distribution of $b_1$

The distribution of slope estimates $b_1$, across all different samples

```{r, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE, fig.height=1.5}
ggplot(data = flights, mapping = aes(x = distance,  y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Population Intercept: ", beta[0], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[1], digits = 4)), ", Population Slope: ", beta[1], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[2], digits = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height=1.5, warning = FALSE, eval = TRUE}
n <- 20
flights_sample <- flights %>% sample_n(size = n)
sample_lm <- lm(air_time ~ distance, data = flights_sample)
sample_slopes <- bind_rows(
  sample_slopes,
  data.frame(
    sample_slope = coef(sample_lm)[2]
  )
)

ggplot(data = flights_sample, mapping = aes(x = distance, y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Sample Intercept: ", b[0], " = ", .(format(round(coef(sample_lm)[1], digits = 4), nsmall = 4)), ", Sample Slope: ", b[1], " = ", .(format(round(coef(sample_lm)[2], digits = 4), nsmall = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height = 1.25, eval = TRUE}
ggplot() +
  geom_histogram(mapping = aes(x = sample_slope, y = ..density..), binwidth = 0.001, data = sample_slopes) +
  xlab("Sample Slopes, n = 20") +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm")) +
  xlim(c(0.1, 0.16))
```

## Sampling Distribution of $b_1$

The distribution of slope estimates $b_1$, across all different samples

```{r, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE, fig.height=1.5}
ggplot(data = flights, mapping = aes(x = distance,  y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Population Intercept: ", beta[0], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[1], digits = 4)), ", Population Slope: ", beta[1], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[2], digits = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height=1.5, warning = FALSE, eval = TRUE}
n <- 20
flights_sample <- flights %>% sample_n(size = n)
sample_lm <- lm(air_time ~ distance, data = flights_sample)
sample_slopes <- bind_rows(
  sample_slopes,
  data.frame(
    sample_slope = coef(sample_lm)[2]
  )
)

ggplot(data = flights_sample, mapping = aes(x = distance, y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Sample Intercept: ", b[0], " = ", .(format(round(coef(sample_lm)[1], digits = 4), nsmall = 4)), ", Sample Slope: ", b[1], " = ", .(format(round(coef(sample_lm)[2], digits = 4), nsmall = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height = 1.25, eval = TRUE}
ggplot() +
  geom_histogram(mapping = aes(x = sample_slope, y = ..density..), binwidth = 0.001, data = sample_slopes) +
  xlab("Sample Slopes, n = 20") +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm")) +
  xlim(c(0.1, 0.16))
```

## Sampling Distribution of $b_1$

The distribution of slope estimates $b_1$, across all different samples

```{r, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE, fig.height=1.5}
ggplot(data = flights, mapping = aes(x = distance,  y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Population Intercept: ", beta[0], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[1], digits = 4)), ", Population Slope: ", beta[1], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[2], digits = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height=1.5, warning = FALSE, eval = TRUE}
n <- 20
flights_sample <- flights %>% sample_n(size = n)
sample_lm <- lm(air_time ~ distance, data = flights_sample)
sample_slopes <- bind_rows(
  sample_slopes,
  data.frame(
    sample_slope = coef(sample_lm)[2]
  )
)

ggplot(data = flights_sample, mapping = aes(x = distance, y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Sample Intercept: ", b[0], " = ", .(format(round(coef(sample_lm)[1], digits = 4), nsmall = 4)), ", Sample Slope: ", b[1], " = ", .(format(round(coef(sample_lm)[2], digits = 4), nsmall = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height = 1.25, eval = TRUE}
ggplot() +
  geom_histogram(mapping = aes(x = sample_slope, y = ..density..), binwidth = 0.001, data = sample_slopes) +
  xlab("Sample Slopes, n = 20") +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm")) +
  xlim(c(0.1, 0.16))
```


## Sampling Distribution of $b_1$

The distribution of slope estimates $b_1$, across all different samples

```{r, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE, fig.height=1.5}
ggplot(data = flights, mapping = aes(x = distance,  y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Population Intercept: ", beta[0], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[1], digits = 4)), ", Population Slope: ", beta[1], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[2], digits = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height=1.5, warning = FALSE, eval = TRUE}
n <- 20
flights_sample <- flights %>% sample_n(size = n)
sample_lm <- lm(air_time ~ distance, data = flights_sample)
sample_slopes <- bind_rows(
  sample_slopes,
  data.frame(
    sample_slope = coef(sample_lm)[2]
  )
)

ggplot(data = flights_sample, mapping = aes(x = distance, y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Sample Intercept: ", b[0], " = ", .(format(round(coef(sample_lm)[1], digits = 4), nsmall = 4)), ", Sample Slope: ", b[1], " = ", .(format(round(coef(sample_lm)[2], digits = 4), nsmall = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height = 1.25, eval = TRUE}
ggplot() +
  geom_histogram(mapping = aes(x = sample_slope, y = ..density..), binwidth = 0.001, data = sample_slopes) +
  xlab("Sample Slopes, n = 20") +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm")) +
  xlim(c(0.1, 0.16))
```




```{r, echo = FALSE, fig.height=1.25}
sample_slopes <- bind_rows(
  sample_slopes,
  {do(10000) * {
    flights_sample <- flights %>% sample_n(size = n)
    sample_lm <- lm(air_time ~ distance, data = flights_sample)
    data.frame(
      sample_slope = coef(sample_lm)[2]
    )
  }} %>% select(sample_slope)
)
```

## Sampling Distribution of $b_1$

The distribution of slope estimates $b_1$, across all different samples

```{r, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE, fig.height=1.5}
ggplot(data = flights, mapping = aes(x = distance,  y = air_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  ggtitle(bquote(paste("Population Intercept: ", beta[0], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[1], digits = 4)), ", Population Slope: ", beta[1], " = ", .(round(coef(lm(air_time ~ distance, data = flights))[2], digits = 4))))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height=1.5, warning = FALSE, eval = TRUE}
ggplot(data = flights_sample, mapping = aes(x = distance, y = air_time)) +
  xlim(c(0, 3000)) +
  ylim(c(0, 600)) +
  geom_text(mapping = aes(x = x, y = y, label = label), size = 8,
    data = data.frame(x = 1500, y = 300, label = "( ...draw 10,000 more samples of size n... )")) +
  ggtitle(bquote(paste("Sample Intercept: ", b[0], " = ...", ", Sample Slope: ", b[1], " = ..."))) +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm"))
```

```{r, echo = FALSE, fig.height = 1.25, eval = TRUE, warning = FALSE}
ggplot() +
  geom_histogram(mapping = aes(x = sample_slope, y = ..density..), binwidth = 0.001, data = sample_slopes) +
  xlab("Sample Slopes, n = 20") +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm")) +
  xlim(c(0.1, 0.16))
```

## Sampling Distribution of $b_1$

* If all of the conditions for inference are satisfied (R. O'LINE) then

$b_1 \sim \text{Normal}\left(\beta_1, SD(b_1)\right)$, where $SD(b_1) = \sqrt{\frac{\sigma^2}{\sum_{i=1}^n (x_i - \bar{x})^2 }}$

```{r, echo = FALSE, fig.height = 1.25, eval = TRUE, warning = FALSE}
ggplot(data = sample_slopes, mapping = aes(x = sample_slope)) +
  geom_histogram(mapping = aes(y = ..density..), binwidth = 0.001) +
  stat_function(
    fun = dnorm,
    args = list(mean = coef(pop_lm)[2], sd = sqrt(var(pop_resids) / sum((flights_sample$distance - mean(flights_sample$distance))^2))),
    color = "red"
  ) +
  xlab("Sample Slopes, n = 20") +
  theme(plot.margin = unit(x = c(0, 0, 0, 0), units = "cm")) +
  xlim(c(0.1, 0.16)) #    args = list(mean = mean(sample_slopes$sample_slope), sd = sd(sample_slopes$sample_slope)),
```

* (This is also still approximately true if most of the assumptions are mostly satisfied.)

* Recall: Probabilities involving the normal distributions only depend on how many standard deviations away from the mean we are:

$\frac{b_1 -\beta_1}{SD(b_1)}$

* **Problem**: This is not useful in practice, because we do not know $\sigma$ (actual standard deviation of residuals in the population), so can't find $SD(b_1)$

## What can we do?

* Estimate $SD(b_1)$.  An estimate of a standard deviation is called a standard error.

$SE(b_1) = \left[\frac{\sum_{i=1}^n (y_i - \hat{y}_i)^2}{(n - 2)\sum_{i = 1}^n (x_i - \bar{x})^2}\right]^{1/2}$

$\frac{b_1 - \beta_1}{SE(b_1)} \sim t_{n-2}$

## How to use this for hypothesis tests?

Null hypothesis: $\beta_1 = 0$

Alternative hypothesis: $\beta_1 \neq 0$

 * **p-value:** Probability of getting a test statistic at least as extreme as what we got based on this sample, assuming the null hypothesis is true.
 
 * **test statistic:** $t = \frac{b_1 - \beta_1}{SE(b_1)} \sim t_{n-2}$
 
 * If null hypothesis is true, $t = \frac{b_1 - \beta_1}{SE(b_1)} = \frac{b_1 - 0}{SE(b_1)}$
    * "How many estimated standard deviations away from the hypothesized slope was our sample slope?"

 * (Calculation of p-value hand-drawn on board)

## How to use this for Conf. Intervals?

 * For a 95% CI, find the value $t^*$ with $P(-t^* \leq \frac{b_1 - \beta_1}{SE(b_1)} \leq t^*) = 0.95$

 * This means that for 95% of samples, $-t^* \leq \frac{b_1 - \beta_1}{SE(b_1)} \leq t^*$
 
 * ...so for 95% of samples, $-t^* SE(b_1) \leq b_1 - \beta_1 \leq t^* SE(b_1)$
 
 * ...so for 95% of samples, $-b_1 -t^* SE(b_1) \leq - \beta_1 \leq -b_1 + t^* SE(b_1)$

 * ...so for 95% of samples, $b_1 -t^* SE(b_1) \leq \beta_1 \leq b_1 + t^* SE(b_1)$

 * Confidence interval: $[b_1 -t^* SE(b_1), b_1 + t^* SE(b_1)]$
 
 * In R, $t^*$ = `qt(0.975, df = n - 2)` for a 95% CI.
 