---
title: 'Stat 140: Regression Summary'
output:
  pdf_document:
    fig_height: 2.7
    fig_width: 6.5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
geometry: margin=1.5cm
header-includes:
   - \usepackage{multicol}
---


```{r, include=FALSE}
require(readr)
require(tidyr)
require(dplyr)
require(ggplot2)
require(knitr)
require(mvtnorm)

options(digits=4)

opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

\vspace{-0.5cm}

#### Population Model

 * The relationship between explanatory and response variable in the population is described by a line with intercept $\beta_0$ and slope $\beta_1$, with normally distributed "errors" around the line
\vspace{-0.5cm}
\begin{align*}
Y_i = \beta_0 + \beta_1 X_i + \varepsilon_i \\
\varepsilon_i \sim \text{Normal}(0, \sigma)
\end{align*}

 * $Y_i$ is response variable value for observational unit number $i$ (number of foals in $i$'th herd)
 * $X_i$ is explanatory variable value for observational unit number $i$ (number of adults in $i$'th herd)
 * $\beta_0$ and $\beta_1$ are **population parameters** we want to estimate

\vspace{-0.5cm}

#### Plot line based on sample data

```{r, echo = FALSE, message = FALSE}
horses <- read_csv("https://mhc-stat140-2017.github.io/data/sdm4/Wild_Horses.csv")
```

```{r, echo = TRUE, fig.height=2}
ggplot(data = horses, mapping = aes(x = Adults, y = Foals)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

\vspace{-0.5cm}

#### Fit linear regression model, print summary (`Foals` is response, `Adults` is explanatory)

```{r, echo = FALSE, eval = TRUE}
lm_fit <- lm(Foals ~ Adults, data = horses)
```

```{r, echo = TRUE, eval = FALSE}
lm_fit <- lm(Foals ~ Adults, data = horses)
summary(lm_fit)
```

\includegraphics[height=3in]{R_output.png}

\vspace{-0.5cm}

#### Conditions for inference

**R**epresentative sample; No **O**utliers; **L**inear relationship; **I**ndependent observations; **N**ormally distributed residuals; **E**qual variance of residuals

\newpage

\vspace{-0.5cm}
#### Equation of estimated line based on sample

Predicted Foals = -1.578 + 0.154 $\times$ Adults

\vspace{-0.5cm}

#### Interpretation of Estimated Intercept

The model predicts that if a herd contains 0 Adults, there will be -1.578 Foals born.

\vspace{-0.5cm}

#### Interpretation of Estimated Slope

The model predicts that for each additional Adult in a herd of horses, an additional 0.154 Foals will be born.

\vspace{-0.5cm}

#### Prediction for a Herd with 50 Adults

Predicted Foals = -1.578 + 0.154 $\times$ 50 = 6.122.

The model predicts that a herd with 50 adults will have 6.122 foals.

\vspace{-0.5cm}

#### Find and interpret a 95% confidence interval for $\beta_1$ (procedure similar for $\beta_0$)

Confidence Interval for $\beta_1$: $b_1 \pm t^* SE(b_1)$, where:

 * $b_1$ is estimate of slope based on this sample (from the R summary output)
 * $t^*$ is the critical value from a $t$ distribution with $n - 2$ degrees of freedom (from `qt`)
 * $SE(b_1)$ is the standard error of $b_1$ (from the R summary output)

```{r}
confint(lm_fit, level = 0.95)
0.154 - qt(0.975, df = 36) * 0.0114
0.154 + qt(0.975, df = 36) * 0.0114
```

We are 95% confident that the slope of a line describing the relationship between the number of adults in a herd of horses and the number of foals born to that herd, in the population of all herds of horses, is between 0.13 and 0.18.

\vspace{-0.5cm}

#### Conduct a hypothesis test with null hypothesis $\beta_1 = 0$

Test statistic: $t = \frac{b_1 - \beta_1^{null}}{SE(b_1)} \sim t_{n - 2}$

```{r}
(0.154 - 0)/0.0114
2 * pt(-13.5, df = 36)
```

Note that the third column of the "Coefficients:" table on the previous page also has the test statistic for this test, and the fourth column has the p-value.  The notation $1.175e{-}15$ means $1.175 * 10^{-15} = 0.000000000000001175$.  Since the p-value is very small, we reject the null hypothesis.  The data provide strong evidence that there is an association between the number of adults in a herd and the number of foals born to that herd.

\vspace{-0.5cm}

#### Use the residual standard deviation to describe how good the model's predictions are.

About 95% of predictions from this model are within plus or minus 9.88 foals of the actual number of foals produced by a herd.  (9.88 is two times the residual standard deviation from the R summary output.)

\vspace{-0.5cm}

#### Use the $R^2$ value to describe how useful the model is (not that important, included for completeness)

The $R^2$ value for this regression is 0.835.  This is close to 1, indicating that the points fall fairly close to the line.  (Recall that $R^2$ is the square of the correlation between the explanatory and response variables.)  This linear model accounts for about 83.5% of the variation in the response variable.